---
title: "CFA"
author: "Elena Shaw"
date: "18/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(lavaan)
library(foreign)
library(dplyr)
library(ggplot2)

setwd('/Users/elenashaw/Documents/UoE/Dissertation/BilingualChildren/code/BilingualChildren/')

MI_data = read.csv('data/MI_complete_data.csv') %>%
# MI_data = read.csv('data/MI_complete_out_rm.csv') %>%
            mutate(BIH = brief_raw_inhibit,
                    BSM = brief_raw_self.monitor,
                    BS = brief_raw_shift,
                    BEC = brief_raw_emotional_control,
                    BIT = brief_raw_initiate,
                    BWM = brief_raw_working_memory,
                    BPO = brief_raw_plan_organise,
                    BTM = brief_raw_task_monitor,
                    BOM = brief_raw_organisation_of_materials,
                    FCE = flanker_percenterrors_congruent,
                    FIE = flanker_percenterrors_incongruent,
                    FCR = flanker_mean_rt_congruent,
                    FIR = flanker_mean_rt_incongruent,
                    PBR = pvt_mean_rt,
                    PLC = pvt_number_of_lapses,
                    PLR = pvt_mean_lapse_rt,
                    PFC = pvt_count_falsestarts
            )

vars_to_transform = c('PBR','PLR') #select(contains(c("B_","rt","ct","err")))

# B_init = brief_raw_initiate,
#  B_inhi = brief_raw_inhibit,
#  B_sm = brief_raw_self.monitor,
#  B_s = brief_raw_shift,
#  B_ec = brief_raw_emotional_control,
#  B_wm = brief_raw_working_memory,
#  B_po = brief_raw_plan_organise,
#  B_tm = brief_raw_task_monitor,
#  B_oom = brief_raw_organisation_of_materials,
#  con_err = flanker_percenterrors_congruent,
#  con_rt = flanker_mean_rt_congruent,
#  inc_err = flanker_percenterrors_incongruent,
#  inc_rt = flanker_mean_rt_incongruent,
#  rt = pvt_mean_rt,
#  fs_ct = pvt_count_falsestarts,
#  lap_rt = pvt_mean_lapse_rt,
#  lap_ct = pvt_number_of_lapses,
```
marker method: fixes the first loading of each factor to 1
variance standardization method: fixes the variance of each factor to 1 but freely estimates all loadings.
## Single factor loadings
```{r}
MI_std = MI_data %>% select(contains(c("B_","rt","ct","err")))%>%
              mutate_at(PBR = log(PBR),
                        PLR = log(PLR))

factor1 = 'f1 =~ BPO + BIT + BTM + BSM + BS + BWM' # + BIH' # + BEC + BOM'
cfa_f1_marker = cfa.mi(factor1, data=MI_std, m=5)
summary(cfa_f1_marker,standardized=TRUE)
cfa_f1_std = cfa.mi(factor1, data=MI_std, m=5, std.lv=TRUE)
summary(cfa_f1_std,standardized=TRUE, fit.measures=TRUE)

inspect(cfa_f1_std, "fit")
inspect(cfa_f1_std, "impute")

lavTestLRT(cfa_f1_std)

factor2 = 'f2 =~ con_err + con_rt + rt + lap_rt' #+ inc_err + inc_rt
cfa_f2_marker = cfa.mi(factor2, data=MI_std, m=5)
summary(cfa_f2_marker,standardized=TRUE)
cfa_f2_std = cfa.mi(factor2, data=MI_std, m=5, std.lv=TRUE)
summary(cfa_f2_std,standardized=TRUE, fit.measures=TRUE)
```


Conclusions:
1) Outlier removed
- none are shown to have good fit; only model which achieves "good approximation" is single factor on BRIEF with only B_po, B_init, B_tm, B_sm, B_s, B_wm, and MAYBE B_oom.
- single factor has THE worst fit across all reported metrics
- fit on non-brief items is not good; fit is maximized with all indicators except false starts; it is further improved if lapse count is also removed.

2) Full dataset
- Both show perfect/good fit (RMSE = 0.00, CFI=1, F-test > 50%)
- brief has better overall fit than flanker+pvt (except by AIC/BIC)

## consider 2 factors
```{r}
#uncorrelated two factor solution, var std method
uncorr1_f2eq ='f1 =~ B_po + B_init + B_tm + B_sm + B_s + B_wm
          f2 =~ a*(con_err + con_rt + rt + lap_rt)
          f1 ~~ 0*f2 ' 
uncorr1_f2eq = cfa.mi(uncorr1_f2eq, data = MI_std, m=5, std.lv=TRUE) 
summary(uncorr1_f2eq, fit.measures=TRUE, standardized=TRUE)

uncorr2_f2eq ='f1 =~ a*(B_po + B_init + B_tm + B_sm + B_s + B_wm)
          f2 =~ con_err + con_rt + rt + lap_rt
          f1 ~~ 0*f2 ' 
uncorr2_f2eq = cfa.mi(uncorr2_f2eq, data = MI_std, m=5, std.lv=TRUE) 
summary(uncorr2_f2eq, fit.measures=TRUE, standardized=TRUE)

# BAD FIT:
# uncorr_f1eq ='f1 =~ a*(B_po + B_init + B_tm + B_inhi + B_sm + B_s + B_ec + B_wm + B_oom)
#           f2 =~ con_err + con_rt + inc_err + inc_rt + rt + lap_rt + lap_ct
#           f1 ~~ 0*f2 ' 
# uncorr_f1eq = cfa(uncorr_f1eq, data = MI_std, std.lv=TRUE) 
# summary(uncorr_f1eq, fit.measures=TRUE, standardized=TRUE)

#correlated two factor solution, marker method
corr1 = 'f1 =~ B_po + B_init + B_tm + B_sm + B_s + B_wm
        f2 =~ con_err + con_rt + rt + lap_rt' 
# B_po + B_init + B_tm + B_inhi + B_sm + B_s + B_ec + B_wm + B_oom
# con_err + con_rt + inc_err + inc_rt + rt + lap_rt
corr1_2factors = cfa.mi(corr1, data = MI_std, m=5, std.lv=TRUE) 
summary(corr_2factors, fit.measures=TRUE, standardized=TRUE)

# good fit, but unjustified removal for the increase in fit
# corr2 = 'f1 =~ B_po + B_init + B_tm + B_sm + B_wm
#         f2 =~ con_err + con_rt + rt + lap_rt' 
# corr2_2factors = cfa(corr2, data = MI_std, std.lv=TRUE) 
# summary(corr2_2factors, fit.measures=TRUE, standardized=TRUE)
```

Conclusions:
1) Outliers removed
- none are shown to have good fit.
- uncorrelated factors with factor 2 set to have equal contribution fits best
- if correlated, removing lapse count improves fit

2) Full dataset
- Reasonable fit using combinations determined above (on individual factors)
- Factors are likely uncorrelated (worst fit and low correlation, -.25)
- Uncorr1 has near perfect fit by RMSE, but uncorr2 has lower AIC/BIC

## 2nd-order?
```{r}
#+ con_rt + inc_rt
mix_factor = 'f1 =~ B_po + B_init + B_tm + B_sm + B_s + B_wm
                f2 =~ con_err + con_rt + rt + lap_rt
                f3 =~ 1*f1 + 1*f2
                f3 ~~ f3 '
cfa_mf_std = cfa.mi(mix_factor, data=MI_std, m=5, std.lv=TRUE)
summary(cfa_mf_std,standardized=TRUE, fit.measures=TRUE)

```