---
title: "EFA"
author: "Elena Shaw"
date: "10/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(psych)
library(dplyr)
library(ggplot2)

# MI_corr = read.csv('data/MI_max_corrs.csv')
MI_corr = read.csv('data/MI_cons_corrs.csv')
MI1_corr = MI_corr %>% dplyr::filter(imp==1) %>% select(-imp)
MI2_corr = MI_corr %>% dplyr::filter(imp==2) %>% select(-imp)
MI3_corr = MI_corr %>% dplyr::filter(imp==3) %>% select(-imp)
MI4_corr = MI_corr %>% dplyr::filter(imp==4) %>% select(-imp)
MI5_corr = MI_corr %>% dplyr::filter(imp==5) %>% select(-imp)
```

## Number of factors? PCA
```{r}
data1_pca1 = principal(MI1_corr,nfactors = 1, rotate = 'none')
data1_pca2 = principal(MI1_corr,nfactors = 2, rotate = 'none')
# data1_pca1$communality
# data1_pca2$communality

principal(MI1_corr,nfactors = 1, rotate = 'none')
principal(MI1_corr,nfactors = 2, rotate = 'none')

principal(MI2_corr,nfactors = 1, rotate = 'none')
principal(MI2_corr,nfactors = 2, rotate = 'none')
principal(MI2_corr,nfactors = 3, rotate = 'none')

principal(MI3_corr,nfactors = 1, rotate = 'none')
principal(MI3_corr,nfactors = 2, rotate = 'none')
principal(MI3_corr,nfactors = 3, rotate = 'none')

principal(MI4_corr,nfactors = 1, rotate = 'none')
principal(MI4_corr,nfactors = 2, rotate = 'none')
principal(MI4_corr,nfactors = 3, rotate = 'none')

principal(MI5_corr,nfactors = 1, rotate = 'none')
principal(MI5_corr,nfactors = 2, rotate = 'none')
principal(MI5_corr,nfactors = 3, rotate = 'none')
```

## Number of factors? Parallel Analysis
```{r}
data1_fa1 = fa.parallel(MI1_corr, nfactors = 1, fm="pa", n.iter=100)
data1_fa2 = fa.parallel(MI1_corr, nfactors = 2, fm="pa", n.iter=100)

# data1_fa2$fa.values
# data1_fa$e.values
# plot(data1_fa)

data2_fa1 = fa.parallel(MI2_corr, nfactors = 1, fm="pa", n.iter=100)
data2_fa2 = fa.parallel(MI2_corr, nfactors = 2, fm="pa", n.iter=100)

fa.parallel(MI3_corr, nfactors = 1, fm="pa", n.iter=100)
fa.parallel(MI3_corr, nfactors = 2, fm="pa", n.iter=100)
fa.parallel(MI4_corr, nfactors = 1, fm="pa", n.iter=100)
fa.parallel(MI4_corr, nfactors = 2, fm="pa", n.iter=100)
fa.parallel(MI5_corr, nfactors = 1, fm="pa", n.iter=100)
fa.parallel(MI5_corr, nfactors = 2, fm="pa", n.iter=100)
```
## Number of factors? Simple Structure
```{r}
# orthoganol rotation
vss1 = vss(MI1_corr,n=5, n.obs=89)
vss1_pa = vss(MI1_corr,n=5, fm="pa", n.obs=89)

vss2 = vss(MI2_corr,n=5, n.obs=89)
vss2_pa = vss(MI2_corr,n=5, fm="pa", n.obs=89)

vss3 = vss(MI3_corr,n=5, n.obs=89)
vss3_pa = vss(MI3_corr,n=5, fm="pa", n.obs=89)

vss4 = vss(MI4_corr,n=5, n.obs=89)
vss4_pa = vss(MI4_corr,n=5, fm="pa", n.obs=89)

vss5 = vss(MI5_corr,n=5, n.obs=89)
vss5_pa = vss(MI5_corr,n=5, fm="pa", n.obs=89)

summary(vss1)
summary(vss1_pa)
summary(vss2)
summary(vss2_pa)
summary(vss3)
summary(vss3_pa)
summary(vss4)
summary(vss4_pa)
summary(vss5)
summary(vss5_pa)

# oblique rotation
vss1_ob = vss(MI1_corr,n=5, rotate = "oblimin", n.obs=89)
vss1_pa_ob = vss(MI1_corr,n=5, rotate = "oblimin", fm="pa", n.obs=89)

vss2_ob = vss(MI2_corr,n=5, rotate = "oblimin", n.obs=89)
vss2_pa_ob = vss(MI2_corr,n=5, rotate = "oblimin", fm="pa", n.obs=89)

vss3_ob = vss(MI3_corr,n=5, rotate = "oblimin", n.obs=89)
vss3_pa_ob = vss(MI3_corr,n=5, rotate = "oblimin", fm="pa", n.obs=89)

vss4_ob = vss(MI4_corr,n=5, rotate = "oblimin", n.obs=89)
vss4_pa_ob = vss(MI4_corr,n=5, rotate = "oblimin", fm="pa", n.obs=89)

vss5_ob = vss(MI5_corr,n=5, rotate = "oblimin", n.obs=89)
vss5_pa_ob = vss(MI5_corr,n=5, rotate = "oblimin", fm="pa", n.obs=89)

summary(vss1_ob)
summary(vss1_pa_ob)
summary(vss2_ob)
summary(vss2_pa_ob)
summary(vss3_ob)
summary(vss3_pa_ob)
summary(vss4_ob)
summary(vss4_pa_ob)
summary(vss5_ob)
summary(vss5_pa_ob)
```
## Number of factors? Clustering
```{r}
# pdf(file='plots/iclust_max_corr.pdf', width=16)
iclust(as.matrix(MI1_corr))
# dev.off()

iclust(as.matrix(MI2_corr))
iclust(as.matrix(MI3_corr))
iclust(as.matrix(MI4_corr))
iclust(as.matrix(MI5_corr))
```

## EFA
```{r}
# oblique rotation
fa1_MR_output = fa(MI1_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
fa1_PA_output = fa(MI1_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')

fa1_MR_output
fa1_PA_output

fa(MI2_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
fa(MI2_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')

fa(MI3_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
fa(MI3_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')

fa(MI4_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
fa(MI4_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')

fa(MI5_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
fa(MI5_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')
```
For conservative correlation, there is some desire to split up and create a 3rd group, but this is usually only made up of 1 indicator. Depending on which imputed data is used, that extra group would be made up of only "pvt lapses/false starts" or the "flanker incongruent".

Factor Analysis REALLY does not like the ovt_count_falsestarts

## Factor Interpretation: 2 fators
```{r}
unrotated1 = fa(MI1_corr,nfactors=2,n.obs=89, rotate='none', n.iter=100, max.iter=1000)
rotated1 = fa(MI1_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
# fa(MI1_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')

unrotated1$loadings
rotated1$loadings
```
```{r}
fa(MI1_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI2_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI3_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI4_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI5_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)$loadings
```
## Factor Interpretation: 3 fators
```{r}
fa(MI1_corr,nfactors=3,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI2_corr,nfactors=3,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI3_corr,nfactors=3,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI4_corr,nfactors=3,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI5_corr,nfactors=3,n.obs=89, n.iter=100, max.iter=1000)$loadings
```

IM_data1: 3rd category: pvt_count_falsestarts (.871) + maybe pvt_number_of_lapses (-.345)
IM_data2: 3rd category: flanker_percenterrors_congruent (.590) + flanker_percenterrors_incongruent (.713) + flanker_mean_rt_incongruent (.521
)
IM_data3: 3rd cateogry: flanker_percenterrors_congruent (.527) + pvt_count_falsestarts (.412)
IM_data4: 3rd: all flanker indicators + no pvt_count_falsestarts
IM_data5: 3rd: flanker_mean_rt_congruent (.809) + flanker_mean_rt_incongruent (.469) + maybe pvt_number_of_lapses (.327)


## Conclusion
Either correlation ultimately does better with 2 factors. The liberal correlation makes this very clear with minimal disagreement among the methods. This is less clear with the conservative analysis. There is a desire to create a 3rd factor but what that 3rd factor would encompass is not stable. It appears to want to split up the 3 indicator types, but the cross-loading on the flanker and pvt indicators tend to be fairly equal.
