---
title: "EFA"
author: "Elena Shaw"
date: "10/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(psych)
library(dplyr)
library(ggplot2)

# MI_corr = read.csv('data/MI_max_corrs.csv')
MI_corr = read.csv('data/MI_cons_corrs.csv')
MI1_corr = MI_corr %>% dplyr::filter(imp==1) %>% select(-imp)
MI2_corr = MI_corr %>% dplyr::filter(imp==2) %>% select(-imp)
MI3_corr = MI_corr %>% dplyr::filter(imp==3) %>% select(-imp)
MI4_corr = MI_corr %>% dplyr::filter(imp==4) %>% select(-imp)
MI5_corr = MI_corr %>% dplyr::filter(imp==5) %>% select(-imp)

```

## Number of factors? PCA
```{r}
eigen(MI_corr_full)
eigen(MI_corr_rm)

pca_full = principal(MI_corr_full,nfactors = 3, rotate = 'none', n.obs=89)
pca_rm = principal(MI_corr_rm,nfactors = 3, rotate = 'none', n.obs=89)

Evec_full = eigen(MI_corr_full)$vectors[,1:3]
rownames(Evec_full) = colnames(MI_corr_full)

Evec_rm = eigen(MI_corr_rm)$vectors[,1:3]
rownames(Evec_rm) = colnames(MI_corr_rm)

# saveRDS(Evec_full, 'data/cons_Evec_full.rds')
# saveRDS(Evec_rm, 'data/cons_Evec_rm.rds')
```

## Number of factors? Parallel Analysis
```{r}
fa.parallel(MI_corr_full, nfactors = 3, fa="fa", fm="pa", n.iter=1000)
fa.parallel(MI_corr_rm, nfactors = 3, fa="fa", fm="pa", n.iter=1000)
```
## Number of factors? Simple Structure
```{r}
# oblique rotation
vss_ob = vss(MI_corr_full,n=5, rotate = "oblimin", n.obs=89)
vss_pa_ob = vss(MI_corr_full,n=5, rotate = "oblimin", fm="pa", n.obs=89)
summary(vss_ob)
summary(vss_pa_ob)
```
## Number of factors? Clustering
```{r}
# pdf(file='plots/iclust_max_corr.pdf', width=16)
iclust(as.matrix(MI_corr_full))
iclust(as.matrix(MI_corr_rm))
# dev.off()
```

## EFA
```{r}
# oblique rotation
fa(MI_corr_full,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
fa(MI_corr_full,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')

fa(MI_corr_rm,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
fa(MI_corr_rm,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')
```
For conservative correlation, there is some desire to split up and create a 3rd group, but this is usually only made up of 1 indicator. Depending on which imputed data is used, that extra group would be made up of only "pvt lapses/false starts" or the "flanker incongruent".

Factor Analysis REALLY does not like the ovt_count_falsestarts

## Factor Interpretation: 2 fators
```{r}
unrotated = fa(MI_corr_full,nfactors=2,n.obs=89, rotate='none', n.iter=100, max.iter=1000)
rotated = fa(MI_corr_full,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)
# fa(MI1_corr,nfactors=2,n.obs=89, n.iter=100, max.iter=1000, fm='pa')

unrotated$loadings
rotated$loadings
```
```{r}
fa(MI_corr_full,nfactors=2,n.obs=89, n.iter=100, max.iter=1000)$loadings
```
## Factor Interpretation: 3 factors
```{r}
fa(MI_corr_full,nfactors=3,n.obs=89, n.iter=100, max.iter=1000)$loadings
fa(MI_corr_rm,nfactors=3,n.obs=89, n.iter=100, max.iter=1000)$loadings
```

IM_data1: 3rd category: flanker_percenterrors_congruent (.62) + flanker_percenterrors_incongruent (.67) + partial loading for flanker_mean_rt_incongruent (.33)
IM_data2: 3rd category: brief_self.monitor (1.01) + partial initiate (.323) & inhibit (.29)
IM_data3: 3rd cateogry: brief_self.monitor(1.00) + partial initiate (.37) & inhibit (.29)
IM_data4: 3rd: brief_self.monitor(1.00) + partial initiate (.36) & inhibit (.27)
IM_data5: 3rd: brief_self.monitor(.61) + brief_inhibit (.87) + partial initiate (.36) oom (.36)
 
## Conclusion
Either correlation ultimately does better with 2 factors. The liberal correlation makes this very clear with minimal disagreement among the methods. This is less clear with the conservative analysis. There is a desire to create a 3rd factor but what that 3rd factor would encompass is not stable. It appears to want to split up the 3 indicator types, but the cross-loading on the flanker and pvt indicators tend to be fairly equal.

## PCA for language
```{r}
read.csv('data/')
eigen(MI_corr_full)
```
## Conclusion
Either correlation ultimately does better with 2 factors. The liberal correlation makes this very clear with minimal disagreement among the methods. This is less clear with the conservative analysis. There is a desire to create a 3rd factor but what that 3rd factor would encompass is not stable. It appears to want to split up the 3 indicator types, but the cross-loading on the flanker and pvt indicators tend to be fairly equal.
