---
title: "Mediation"
author: "Elena Shaw"
date: "18/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(lavaan)
library(foreign)
library(dplyr)
library(ggplot2)
library(pwr)

setwd('/Users/elenashaw/Documents/UoE/Dissertation/BilingualChildren/code/BilingualChildren/')

demo_data = read.csv('data/data_num.csv')  %>% 
              select(gender, age_m, diagnosis, SCQ,
               bilec_total_input, age_acquisition, where_english) %>%
              mutate(age = age_m/12,
                     years_bilingual = (age - age_acquisition)*bilec_total_input,
                     year_bi_std = scale(years_bilingual)
                     )
MI_full = read.csv('data/MI_complete_data.csv') %>% select(-c(gender,age_m,diagnosis))
# MI_rm = read.csv('data/MI_complete_out_rm.csv') %>% select(-c(gender,age_m,diagnosis))

MI1_full = MI_full %>% filter(.imp==1)
```
## project data (PCA)
```{r}
source("func.R")
EF_Evecs = readRDS('data/cons_Evec_full.rds')
# EF_Evecs = readRDS('data/cons_Evec_rm.rds')
lang_Evecs = readRDS('data/lang_Evec.rds')

proj1_full = project_data(MI1_full,EF_Evecs$data1,lang_Evecs$data1,demo_data)
# proj1_rm = project_data(MI1_rm,EF_Evecs$data1,lang_Evecs$data1,demo_data)


proj2_full = project_data(MI2_full,EF_Evecs$data2,lang_Evecs$data2,demo_data)
# proj2_rm = project_data(MI2_rm,EF_Evecs$data2,lang_Evecs$data2,demo_data)
```

## Moderation:YSB
```{r}
df1 = proj1_full %>%
      tidyr::pivot_longer(c(EF_Comp1,EF_Comp2),names_to="EFComp",values_to="EF") %>%
      mutate(diagnosis = as.factor(diagnosis),
             gender = as.factor(gender),
              mod_bil = EF * years_bilingual,
              mod_age = EF * age,
              )
ggplot(df1, aes(x=diagnosis,y=EF,color=diagnosis)) +
  geom_boxplot() +
  facet_wrap(~ EFComp)
ggplot(df1, aes(x=years_bilingual,y=EF,color=diagnosis)) +
  geom_point() +
  geom_smooth(method = "lm", alpha = .15, aes(fill = diagnosis)) +
  facet_wrap(~ EFComp)
ggplot(df1, aes(x=years_bilingual,y=EF,color=diagnosis)) +
  geom_point() +
  geom_smooth(formula = 'y ~ poly(x,2)',
              method = "lm", alpha = .15, aes(fill = diagnosis)) +
  facet_wrap(~ EFComp)
```
```{r}
proj_data1 = proj1_full
effect_size_diag = cohen.d(proj_data1, "diagnosis")
effect_size_gender = cohen.d(proj_data1, "gender")

t.test(proj_data1$EF_Comp1 ~ proj_data1$diagnosis)
pwr.t.test(n = 89, 
           d = effect_size_diag$cohen.d["EF_Comp1","effect"], 
           sig.level = 0.05)

t.test(proj_data1$EF_Comp2 ~ proj_data1$diagnosis)
pwr.t.test(n = 89, 
           d = effect_size_diag$cohen.d["EF_Comp2","effect"], 
           sig.level = 0.05)
```
Conclusion:
- means are meaningfully difference for EF1 (power = 93.7%), so modeling for EF1 should have separate intercepts
- means are not meaningfully difference for EF2
```{r}
source('func.R')
proj_data1 = proj1_full

# # Does centering make it sig?
# ysb_mo0 = lm(proj_data1$EF_Comp2 ~ proj_data1$diagnosis)
# summary(ysb_mo0)
# ysb_mo1 = lm(proj_data1$EF_Comp2 ~ proj_data1$diagnosis + proj_data1$diagnosis:proj_data1$year_bi_std)
# summary(ysb_mo1)
# ysb_mo2 = lm(proj_data1$EF_Comp2 ~ proj_data1$diagnosis + proj_data1$diagnosis:proj_data1$years_bilingual)
# summary(ysb_mo2)

# EF1
ysb_mo1 = lm(proj_data1$EF_Comp1 ~ proj_data1$diagnosis + poly(proj_data1$years_bilingual,2):proj_data1$diagnosis)
summary(ysb_mo1)
calc_power(ysb_mo1,.05)$power

ysb_mo2 = lm(proj_data1$EF_Comp1 ~ proj_data1$diagnosis + proj_data1$years_bilingual:proj_data1$diagnosis)
summary(ysb_mo2)
calc_power(ysb_mo2,.05)$power

# EF2
ysb_mo3 = lm(proj_data1$EF_Comp2 ~ proj_data1$diagnosis + poly(proj_data1$years_bilingual,2):proj_data1$diagnosis)
summary(ysb_mo3)
calc_power(ysb_mo3,.05)$power

ysb_mo4 = lm(proj_data1$EF_Comp2 ~ proj_data1$diagnosis + proj_data1$years_bilingual:proj_data1$diagnosis)
summary(ysb_mo4)
calc_power(ysb_mo4,.05)$power

# SCQ
TD_data = proj_data1 %>% filter(diagnosis==0)
scq_mo1 = lm(TD_data$EF_Comp2 ~ TD_data$SCQ + TD_data$SCQ:TD_data$years_bilingual)
summary(scq_mo1)
calc_power(scq_mo1,.05)$power
```
## Moderation + gender
```{r}
ggplot(df1, aes(x=diagnosis,y=EF,color=gender)) +
  geom_boxplot() +
  facet_wrap(~ EFComp)
ggplot(df1, aes(x=years_bilingual,y=EF,color=gender)) +
  geom_point() +
  geom_smooth(method = "lm", alpha = .15, aes(fill = gender)) +
  facet_wrap(~ EFComp)
ggplot(df1, aes(x=years_bilingual,y=EF,color=gender)) +
  geom_point() +
  geom_smooth(formula = 'y ~ poly(x,2)',
              method = "lm", alpha = .15, aes(fill = gender)) +
  facet_wrap(~ EFComp)

t.test(proj_data1$EF_Comp1 ~ proj_data1$gender)
pwr.t.test(n = 89, 
           d = effect_size_gender$cohen.d["EF_Comp1","effect"], 
           sig.level = 0.05)

t.test(proj_data1$EF_Comp2 ~ proj_data1$gender)
pwr.t.test(n = 89, 
           d = effect_size_gender$cohen.d["EF_Comp2","effect"], 
           sig.level = 0.05)
```
Conclusion:
- there is no meaningful difference in EF1 or EF2 for girls vs. boys. Therefore, should not have separate intecepts (i.e. gender should not be a covariate)
```{r}

ysb_mo5 = lm(proj_data1$EF_Comp1 ~ proj_data1$diagnosis + proj_data1$diagnosis:proj_data1$gender + proj_data1$years_bilingual:proj_data1$diagnosis)
summary(ysb_mo5)
calc_power(ysb_mo5,.05)$power

ysb_mo6 = lm(proj_data1$EF_Comp2 ~ proj_data1$diagnosis + proj_data1$diagnosis:proj_data1$gender + proj_data1$years_bilingual:proj_data1$diagnosis)
summary(ysb_mo6)
calc_power(ysb_mo6,.05)$power

# Segmentation
# EF1
seg_data1 = proj_data1 %>% filter(gender==0)
seg_mo1 = lm(seg_data1$EF_Comp1 ~ seg_data1$diagnosis + seg_data1$years_bilingual:seg_data1$diagnosis)
summary(seg_mo1)
calc_power(seg_mo1,.05)$power

seg_data2 = proj_data1 %>% filter(gender==1)
seg_mo2 = lm(seg_data2$EF_Comp1 ~ seg_data2$diagnosis + seg_data2$years_bilingual:seg_data2$diagnosis)
summary(seg_mo2)
calc_power(seg_mo2,.05)$power

#EF2
seg_mo3 = lm(seg_data1$EF_Comp2 ~ seg_data1$diagnosis + seg_data1$years_bilingual:seg_data1$diagnosis)
summary(seg_mo3)
calc_power(seg_mo3,.05)$power

seg_mo4 = lm(seg_data2$EF_Comp2 ~ seg_data2$diagnosis + seg_data2$years_bilingual:seg_data2$diagnosis)
summary(seg_mo4)
calc_power(seg_mo4,.05)$power
```
Conclusion:
- No strong evidence that boys and girls differ (too low power)

## Mediation: language skills
```{r}
ggplot(df1, aes(x=diagnosis,y=lang_Comp,color=gender)) +
  geom_boxplot()
ggplot(df1, aes(x=lang_Comp,y=EF,color=gender)) +
  geom_point() +
  geom_smooth(method = "lm", alpha = .15, aes(fill = gender)) +
  facet_wrap(~ EFComp)
ggplot(df1, aes(x=lang_Comp,y=EF,color=gender)) +
  geom_point() +
  geom_smooth(formula = 'y ~ poly(x,2)',
              method = "lm", alpha = .15, aes(fill = gender)) +
  facet_wrap(~ EFComp)


t.test(proj_data1$lang_Comp ~ proj_data1$diagnosis)
pwr.t.test(n = 89, 
           d = effect_size_diag$cohen.d["lang_Comp","effect"], 
           sig.level = 0.05)

t.test(proj_data1$lang_Comp ~ proj_data1$gender)
pwr.t.test(n = 89, 
           d = effect_size_gender$cohen.d["lang_Comp","effect"], 
           sig.level = 0.05)
```
Conlcusion:
- there is no evidence to suggest there is language difference by gender or by diagnosis (the power is also tiny)

## Mediation Model, Step 1:
```{r}
proj_data1 = proj1_full

total_effect1 = lm(proj_data1$EF_Comp1 ~ proj_data1$lang_Comp)
summary(total_effect1)
calc_power(total_effect1,.05)$power

total_effect2 = lm(proj_data1$EF_Comp2 ~ proj_data1$lang_Comp)
summary(total_effect2)
calc_power(total_effect2,.05)$power


# # Moderated?
# t-test suggests not
# total_effect3 = lm(proj_data1$EF_Comp1 ~ proj_data1$lang_Comp +  proj_data1$lang_Comp:proj_data1$diagnosis)
# summary(total_effect3)
# calc_power(total_effect1,.05)$power
# 
# total_effect4 = lm(proj_data1$EF_Comp2 ~ proj_data1$lang_Comp +   proj_data1$lang_Comp:proj_data1$diagnosis)
# summary(total_effect4)
# calc_power(total_effect1,.05)$power

# # Segmentation
# seg_data1 = proj_data1 %>% filter(gender==0)
# total_seg1 = lm(seg_data1$EF_Comp1 ~ seg_data1$lang_Comp)
# summary(total_seg1)
# calc_power(total_seg1,.05)$power
# 
# seg_data2 = proj_data1 %>% filter(gender==1)
# total_seg2 = lm(seg_data2$EF_Comp1 ~ seg_data2$lang_Comp)
# summary(total_seg2)
# calc_power(total_seg2,.05)$power
```

Conclusions:
1) full data set
- EF1 and EF2 somewhat predicted by vocab test (power: ~63%)
  - neither are likely moderated by diagnosis (low power)
  - neither are likely different by gender or diagnosis (low sig. & power)

2) outlier removed
- Brief tests (EF1) well predicted by vocab test (for gender & **ASD**)
  - ** Brief tests (EF1) somewhat predicted by TD vocab test
- Flanker tests (EF2) somewhat predicted by vocab test (for both gender & diagnosis)

- overall moderation by diagnosis? (but NOT moderated mediation)

## Mediation Model, Step 2:
```{r}
proj_data1 = proj1_full

mediator = lm(proj_data1$years_bilingual ~ proj_data1$lang_Comp)
summary(mediator)
calc_power(mediator,.05)$power

# moderated: gender
mediator_gender = lm(proj_data1$years_bilingual ~ proj_data1$lang_Comp + proj_data1$gender:proj_data1$lang_Comp)
summary(mediator_gender)
calc_power(mediator_gender,.05)$power

# moderated: diagnosis
mediator_diag = lm(proj_data1$years_bilingual ~ proj_data1$lang_Comp + proj_data1$diagnosis:proj_data1$lang_Comp)
summary(mediator_diag)
calc_power(mediator_diag,.05)$power

# Segmentation
# seg_data1 = proj_data1 %>% filter(diagnosis==0)
# mediator_seg1 = lm(seg_data1$years_bilingual ~ seg_data1$lang_Comp)
# summary(mediator_seg1)
# calc_power(mediator_seg1,.05)$power
# 
# seg_data2 = proj_data1 %>% filter(diagnosis==1)
# mediator_seg2 = lm(seg_data2$years_bilingual ~ seg_data2$lang_Comp)
# summary(mediator_seg2)
# calc_power(mediator_seg2,.05)$power

proj_data2 = proj2_full
mediator = lm(proj_data2$years_bilingual ~ proj_data2$lang_Comp)
summary(mediator)
calc_power(mediator,.05)$power

proj_data3 = proj3_full
mediator = lm(proj_data3$years_bilingual ~ proj_data3$lang_Comp)
summary(mediator)
calc_power(mediator,.05)$power

proj_data4 = proj4_full
mediator = lm(proj_data4$years_bilingual ~ proj_data4$lang_Comp)
summary(mediator)
calc_power(mediator,.05)$power

proj_data5 = proj5_full
mediator = lm(proj_data5$years_bilingual ~ proj_data5$lang_Comp)
summary(mediator)
calc_power(mediator,.05)$power
```
Conclusions:
1) full data set
- language ability is potentially a predictor of YSB, but low power
  - but unlikely moderated by either diagnosis or gender (power ~48%)
  - not likely different by gender or diagnosis (insig. + low power)
2) outlier removed
- language test is good predictor of bilingualism as measured by YSB but NOT TBI
- ? ASD language performance is more significant than TD in predicting bilingualism
- ** exists gender difference: boys language performance sig. predicts bilingualism
  - (boys language performance remains significant when moderated by age)
  - (boys language performance remains significant when moderated by diagnosis)
- ** Not moderated by gender or diagnosis

## Mediation Model, Step 3 (unmoderated):
```{r}
proj_data1 = proj1_full

dep_var1 = lm(proj_data1$EF_Comp1 ~ proj_data1$lang_Comp + proj_data1$years_bilingual)
summary(dep_var1)
calc_power(dep_var1,.10)$power

dep_var2 = lm(proj_data1$EF_Comp2 ~ proj_data1$lang_Comp + proj_data1$years_bilingual)
summary(dep_var2)
calc_power(dep_var2,.05)$power
```
Conclusion:
1) full data set
- EF1 may be slightly mediated (p = .06, power=.83)
- EF2 may be slightly mediated (p = .04, power=.77/.86)
2) outlier removed
- bilingualism is NOT a good mediator between language and EF performance

## Mediation Model, step 3 (moderated):
```{r}
proj_data1 = proj1_full

mod_med1 = lm(proj_data1$EF_Comp1 ~ proj_data1$lang_Comp + + proj_data1$diagnosis*proj_data1$year_bi_std)
summary(mod_med1)
calc_power(mod_med1,.1)$power

mod_med2 = lm(proj_data1$EF_Comp2 ~ proj_data1$lang_Comp + proj_data1$year_bi_std*proj_data1$diagnosis)
summary(mod_med2)
calc_power(mod_med2,.1)$power

mod_med3 = lm(proj_data1$EF_Comp1 ~ proj_data1$lang_Comp + proj_data1$gender*proj_data1$year_bi_std)
summary(mod_med3)
calc_power(mod_med3,.1)$power

mod_med4 = lm(proj_data1$EF_Comp2 ~ proj_data1$lang_Comp + proj_data1$year_bi_std*proj_data1$gender)
summary(mod_med4)
calc_power(mod_med4,.05)$power

mod_med5 = lm(proj_data1$EF_Comp2 ~ proj_data1$lang_Comp + proj_data1$year_bi_std*proj_data1$diagnosis + proj_data1$year_bi_std*proj_data1$gender)
summary(mod_med5)
calc_power(mod_med5,.1)$power
```
Conclusion:
1) full data set
- diagnosis may be a moderator for std_YSB mediating on EF1 (p:.07, power:82%) and EF2 (p:.07, power: 83%)
- gender may also be moderator for std_YSB mediating on EF1 (p:.05, power:84%) and EF2 (p:.03, power: 79%/87%)

2) outlier removed
- No moderated mediation

## Step 4:
```{r}
med = psych::mediate(EF_Comp1 ~ lang_Comp + (year_bi_std),
              data=proj_data1,
              n.iter = 10000)
summary(med)

mod_med = psych::mediate(EF_Comp1 ~ lang_Comp + diagnosis + (year_bi_std),
              data=proj_data1,
              n.iter = 10000)
summary(mod_med)
```
## Lavaan
```{r}
set.seed(1)

mod_EF1 = "# a path
         year_bi_std ~ a * lang_Comp

         # b path
         EF_Comp1 ~ b * year_bi_std

         # c prime path 
         EF_Comp1 ~ cp * lang_Comp

         # indirect and total effects
         ab := a * b
         total := cp + ab"

mod_EF2 = "# a path
         year_bi_std ~ a * lang_Comp

         # b path
         EF_Comp2 ~ b * year_bi_std

         # c prime path 
         EF_Comp2 ~ cp * lang_Comp

         # indirect and total effects
         ab := a * b
         total := cp + ab"

sem1 = sem(mod_EF1, data = proj_data1, se = "bootstrap", bootstrap = 10000)
sem2 = sem(mod_EF2, data = proj_data1, se = "bootstrap", bootstrap = 10000)

summary(sem1, standardized = T, fit.measures=T)
summary(sem2, standardized = T, fit.measures=T)
# saveRDS(sem1, file='data/mediation_obj1.RDS')
# saveRDS(sem2, file='data/mediation_obj2.RDS')
parameterestimates(sem21)
parameterestimates(sem22)
```

```{r}
set.seed(1)
med_mod1 = "# a path
         year_bi_std ~ a * lang_Comp

         # b path
         EF_Comp1 ~ b1 * year_bi_std + b2 * diagnosis + b3 * diagnosis:year_bi_std

         # c prime path 
         EF_Comp1 ~ cp * lang_Comp

         # indirect and total effects
         ab := a * b1 * b2 *b3
         total := cp + ab"
sem3 = sem(med_mod1, data = proj_data1, se = "bootstrap", bootstrap = 10000)
# saveRDS(sem3, file='data/mod_mediation_obj3.RDS')
# saveRDS(sem3, file='data/mediation_obj3.RDS')
summary(sem3, standardized = T, fit.measures=T)
parameterestimates(sem3)
```